Demo: Counterpoint by Construction

### 2019-06-24 doc/farm/farm-abstract.tex
library overview and future work

TODO: describe the key ingredients of the library.

@@

TODO

%At the FARM workshop, we intend to give a gentle introduction to
%counterpoint, and describe our Agda implementation, showing how the
%type-based approach both aids human composition and allows for 
%computer-generated creation of correct counterpoint.
%We then contrast our work to a recent study on generating
%natural-sounding counterpoint by machine learning
%\citep{CounterpointByConvolution}, which does not provide correctness
%guarantees.
%Finally, we discuss further applications of our library, including
%representation of functional harmony.

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
 The authors would like to thank the participants of the Tokyo Agda 
 Implementors' Meeting, especially Ulf Norell and Jesper Cockx,
 for many helpful suggestions that improved our Agda code.
\end{acks}

>>>>

The goal of Music Tools is to provide a core collection of types and functions
that can be used to easily build applications to analyze and synthesize music.
It is intended to evolved into a dependently-typed replacement for Euterpea
\citep{hudak2018haskell} and borrows many ideas from that library.
It is currently restricted to the chromatic scale for simplicity, and like
Euterpea is designed to work well with MIDI.

The fundamental type for melody is \texttt{Pitch} which is just a natural number with $0$
representing the lowest expressible pitch (it is intended to correspond to MIDI note 0, but
the interpretation can change depending upon the application). One can also
express pitch as a pair of an octave and relative pitch within the ocatve, which is
more convenient for some applications, and convert between this representation and
absolute pitch. One can then prove that converting back and forth is the identity function.

For rhythm the fundamental type is \texttt{Duration}, also a natural number, which represents
some unspecified unit of time (when the music is played the unit is then specified).
A combination of a pitch and a duration gives a \texttt{Note}, and notes in turn are
made into \texttt{Music} via sequential and parallel composition, as in Euterpea.

Music in turn can be converted to MIDI and and played on synthesizers. Existing Haskell
MIDI libraries are called via Agda's Haskell FFI. Details can be found in the github
repository.

@@

We intend to expand the Music Tools library to include
richer functionality for analysis and synthesis. We are
particularly interested in expressing the work done in Haskell
on functional harmony in the library, to see just how much
dependent types can simplify the representation.

Music is a rich yet circumscribed domain in which
issues of equivalence \citep{tabareau2017equivalences}
and ornamentation \citep{dagand2017essence} naturally arise.
For example a
musical score can be interpreted either horizontally (counterpoint) or
vertically (harmony) and it is important to be able to seamlessly
convert between these representations. Similarly one may wish to treat
pitch or rhythm separately as well as combine them, which is a natural
application of ornamentation. We plan to examine the extent to which this
research can be put to use in a practical domain.

For counterpoint specifically, we plan represent higher species counterpoint
and also to explore automatic generation of species counterpoint,
contrasting the music created to recent work on generating
natural-sounding counterpoint by machine learning
\citep{CounterpointByConvolution}, which does not provide correctness
guarantees.

%% Acknowledgments
\begin{acks}                            %% contents suppressed with 'anonymous'
 The authors would like to thank the participants of the Tokyo Agda 
 Implementors' Meeting, especially Ulf Norell and Jesper Cockx,
 for many helpful suggestions that improved our Agda code.
 We also thank the anonymous reviewers for their thoughtful feedback
 which improved our presentation.
\end{acks}
